\begin{acknowledgements}

I feel very fortunate for having the support from many people in my life.
This dissertation would not be possible without them.

First of all, I want to thank my advisor Professor Yiying Zhang for supporting and believing in me during this journey.
In the beginning of my PhD, I had a hard time navigating research and life. I'm very thankful for Yiying's encouragement, trust, and patience during that period.
I'm also very fortunate to work with Yiying on various disaggregation related projects, most of which are somewhat big and grand vision projects.
This type of project is not easy to get through,
but definitely a rewarding experience once done.
Yiying's research taste, critical thinking ability, insights, and attention to detail always amazes me.
Most importantly, she made me realize what type of  a system researcher I want to be and also guided me to become one.

I'm thankful for the researchers and interns during my 2018 and 2019 VMware Research internships.
I thank Stanko Novakovic and Marcos Aguilera for their mentoring and insights.
I also want to thank Sanidhya Kashyap for showing me how knowledgeable and experienced a seasoned system PhD student could be. His dedication and humbleness encouraged me to become a better researcher.
I also thank Yanfang Le, Rohan Kadekodi, Soujanya Ponnapalli, Weili Nie, Arjun Singhvi and many others for that epic internship experience.

I thank Shin-Yeh Tsai, the RDMA guy in our lab, for providing us with a fast and easy to use network stack, also for his suggestions during the start of my PhD.
I thank Yilun Chen, Yutong Huang, Sumukh H. Ravindra for their contribution to the LegoOS project.
I thank Ke Liu for helping out the experiments and plotting during several deadline closings.
I thank Arvind Krishnamurthy's insights and Ryan Kosta's OVS testing in the SuperNIC project.
I thank Will Lin for his contribution in the SuperNIC project,
this project will not be possible without his insightful questions, hard work, commitment, and dedication.
I thank Zhiyuan Guo and Xuhao Luo for their contribution in the Clio project.
I'm very grateful to Alex Forencich and Moein Khazraee for helping me on debugging and bringing up FPGA boards.
The Clio and SuperNIC projects literally would not be possible without their help.

I thank my committee members, Professor Geoffrey M. Voelker, Professor Alex C. Snoeren, Professor Stefan Savage, and Professor George C. Papen for serving as a committee member.
%
I would like to thank Liwei Guo, Heejin Park, Hongyu Miao for the technical discussion back in Purdue. I thank Lixiang Ao, Zachary Blanco, Stewart Grant, Anil Yelam and many others for our discussions in UCSD.

I also want to thank my friends, this journey would not be possible without their love and support. I want to thank Jingwen Zhang, Xiaoyi Huang, Xiaosi Zhu, Heejin Park, Jianyue Zhang, Haolan Liu, Meihan Li, Zesen Zhang, Hanwen Yao, Yi Xu, Jian Yang, Guo Xiang, and many others for making my life in US colorful and meaningful. I wish them nothing but the best.

Finally, I am immensely grateful to my family.
They encouraged me to pursue my dream and provide their unconditional love
through the whole process. There are no words to express my gratitude to my parents, Wenliang Shan and Jinyan Zhang. I dedicate this dissertation to them.

\hfill \break

Chapter 2, in full, is a reprint of Yizhou Shan, Shin-Yeh Tsai, Yiying Zhang, ``Distributed Shared Persistent Memory'', \textit{SoCC, 2017}. The dissertation author was the primary investigator and author of this paper.

Chapter 3, in full, is a reprint of Yizhou Shan, Yutong Huang, Yilun Chen, Yiying Zhang, ``LegoOS: A Disseminated, Distributed OS for Hardware Resource Disaggregation'', \textit{OSDI, 2018}. The dissertation author was the primary investigator and author of this paper.

Chapter 4, in full, is a reprint of Yizhou Shan, Zhiyuan Guo (co-first authors), Xuhao Luo, Yutong Huang, Yiying Zhang, ``Clio: A Hardware-Software Co-Designed Disaggregated Memory System'', \textit{ASPLOS, 2022}. The dissertation author was the primary investigator and author of this paper.

Chapter 5, in part, has been submitted for publication of the material as it may appear in Yizhou Shan, Will Lin, Ryan Kosta, Arvind Krishnamurthy, Yiying Zhang, ``Disaggregating and Consolidating Network Functionalities with SuperNIC'', \textit{arXiv, 2022}. The dissertation author was the primary investigator and author of this paper.

\end{acknowledgements}
